### 2018-05-25 submission/manuscript.Rmd
add references

The bacterial community is reported to vary between case and control for a number of diseases []. However, for diseases like obesity, the taxa identified have varied widely depending on the study []. Some of this variation could be due to error introduced during the 16S rRNA gene sequencing workflow. Yet, standardizing a 16S rRNA gene sequencing workflow will ultimately result in a standardized and reproducible bias due to choices made on the methods used for preservation, extraction, PCR, and sequencing. Within this context, all 16S rRNA gene sequencing methods are biased even when these workflows are standardized to increase reproducibility. In order to interpret specific studies within the broader context of the overall field, assessing error at different parts of the 16S rRNA gene sequencing workflow is critical.

A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, PCR, and sequencing steps. The preservation and extraction stages of the 16S rRNA gene sequencing workflow have been the most extensively studied []. For preservation and extraction stages of the workflow, it has been consistently found that there are biases based on the kits used, but that these differences are smaller than the overall biological difference measured between samples with different kits [@preservation_Song_2016; @storage_Bassis_2017]. Since these studies use the same PCR approach while varying preservation or extraction method, the contribution of PCR bias to this overall workflow is not well characterized.

>>>>

The bacterial community is reported to vary between case and control for a number of diseases [@Turnbaugh2008; @Baxter2016; @Sze2015; @Bonfili2017]. However, for diseases like obesity, the taxa identified have varied widely depending on the study [@Turnbaugh2008; @Zupancic2012]. Some of this variation could be due to error introduced during the 16S rRNA gene sequencing workflow. Yet, standardizing a 16S rRNA gene sequencing workflow will ultimately result in a standardized and reproducible bias due to choices made on the methods used for preservation, extraction, PCR, and sequencing. Within this context, all 16S rRNA gene sequencing methods are biased even when these workflows are standardized to increase reproducibility. In order to interpret specific studies within the broader context of the overall field, assessing error at different parts of the 16S rRNA gene sequencing workflow is critical.

A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, PCR, and sequencing steps. The preservation and extraction stages of the 16S rRNA gene sequencing workflow have been the most extensively studied [@contamination_Salter2014; @preservation_Song_2016; @storage_Bassis_2017; @review_Kim_2017]. For preservation and extraction stages of the workflow, it has been consistently found that there are biases based on the kits used, but that these differences are smaller than the overall biological difference measured between samples with different kits [@preservation_Song_2016; @storage_Bassis_2017]. Since these studies use the same PCR approach while varying preservation or extraction method, the contribution of PCR bias to this overall workflow is not well characterized.

### 2018-05-25 submission/manuscript.tex
add references

The bacterial community is reported to vary between case and control for
a number of diseases {[}{]}. However, for diseases like obesity, the
taxa identified have varied widely depending on the study {[}{]}. Some
of this variation could be due to error introduced during the 16S rRNA
gene sequencing workflow. Yet, standardizing a 16S rRNA gene sequencing
workflow will ultimately result in a standardized and reproducible bias
due to choices made on the methods used for preservation, extraction,
PCR, and sequencing. Within this context, all 16S rRNA gene sequencing
methods are biased even when these workflows are standardized to
increase reproducibility. In order to interpret specific studies within
the broader context of the overall field, assessing error at different
parts of the 16S rRNA gene sequencing workflow is critical.

A typical 16S rRNA gene sequencing workflow can be divided into
preservation, extraction, PCR, and sequencing steps. The preservation
and extraction stages of the 16S rRNA gene sequencing workflow have been
the most extensively studied {[}{]}. For preservation and extraction
stages of the workflow, it has been consistently found that there are
biases based on the kits used, but that these differences are smaller
than the overall biological difference measured between samples with
different kits (Song et al., 2016; Bassis et al., 2017). Since these
studies use the same PCR approach while varying preservation or
extraction method, the contribution of PCR bias to this overall workflow
is not well characterized.

@@

@@

@@

@@

>>>>

The bacterial community is reported to vary between case and control for
a number of diseases (Turnbaugh et al., 2008; Sze et al., 2015; Baxter
et al., 2016; Bonfili et al., 2017). However, for diseases like obesity,
the taxa identified have varied widely depending on the study (Turnbaugh
et al., 2008; Zupancic et al., 2012). Some of this variation could be
due to error introduced during the 16S rRNA gene sequencing workflow.
Yet, standardizing a 16S rRNA gene sequencing workflow will ultimately
result in a standardized and reproducible bias due to choices made on
the methods used for preservation, extraction, PCR, and sequencing.
Within this context, all 16S rRNA gene sequencing methods are biased
even when these workflows are standardized to increase reproducibility.
In order to interpret specific studies within the broader context of the
overall field, assessing error at different parts of the 16S rRNA gene
sequencing workflow is critical.

A typical 16S rRNA gene sequencing workflow can be divided into
preservation, extraction, PCR, and sequencing steps. The preservation
and extraction stages of the 16S rRNA gene sequencing workflow have been
the most extensively studied (Salter et al., 2014; Song et al., 2016;
Bassis et al., 2017; Kim et al., 2017). For preservation and extraction
stages of the workflow, it has been consistently found that there are
biases based on the kits used, but that these differences are smaller
than the overall biological difference measured between samples with
different kits (Song et al., 2016; Bassis et al., 2017). Since these
studies use the same PCR approach while varying preservation or
extraction method, the contribution of PCR bias to this overall workflow
is not well characterized.

@@

\hypertarget{ref-Baxter2016}{}
Baxter NT., Ruffin MT., Rogers MAM., Schloss PD. 2016. Microbiota-based
model improves the sensitivity of fecal immunochemical test for
detecting colonic lesions. \emph{Genome Medicine} 8. DOI:
\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.

@@

\hypertarget{ref-Bonfili2017}{}
Bonfili L., Cecarini V., Berardi S., Scarpona S., Suchodolski JS.,
Nasuti C., Fiorini D., Boarelli MC., Rossi G., Eleuteri AM. 2017.
Microbiota modulation counteracts alzheimer's disease progression
influencing neuronal proteolysis and gut hormones plasma levels.
\emph{Scientific Reports} 7. DOI:
\href{https://doi.org/10.1038/s41598-017-02587-2}{10.1038/s41598-017-02587-2}.

@@

\hypertarget{ref-Turnbaugh2008}{}
Turnbaugh PJ., Hamady M., Yatsunenko T., Cantarel BL., Duncan A., Ley
RE., Sogin ML., Jones WJ., Roe BA., Affourtit JP., Egholm M., Henrissat
B., Heath AC., Knight R., Gordon JI. 2008. A core gut microbiome in
obese and lean twins. \emph{Nature} 457:480--484. DOI:
\href{https://doi.org/10.1038/nature07540}{10.1038/nature07540}.

@@

\hypertarget{ref-Zupancic2012}{}
Zupancic ML., Cantarel BL., Liu Z., Drabek EF., Ryan KA., Cirimotich S.,
Jones C., Knight R., Walters WA., Knights D., Mongodin EF., Horenstein
RB., Mitchell BD., Steinle N., Snitker S., Shuldiner AR., Fraser CM.
2012. Analysis of the gut microbiota in the old order amish and its
relation to the metabolic syndrome. \emph{PLoS ONE} 7:e43052. DOI:
\href{https://doi.org/10.1371/journal.pone.0043052}{10.1371/journal.pone.0043052}.
